---
title: "nwfscSurvey Vignette"
output:
  pdf_document: default
  html_document: default
date: "2025-01-23"
---
Set up: The first set of code is for the Bottom trawl survey
        The second set of code is for the triennial survey

Questions: is this a two area model as in WA/OR + CA? or WA + OR only? Is California a whole different assessment model?
If we are seeing catches in CA, what is the strata doing again?

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(nwfscSurvey)
dir.create(file.path(getwd(),"plots"))
```

## NWFSC.Combo Data

Initial pull and plotting of data: WCGBTS
```{r, warning = FALSE, message = FALSE}
catch = pull_catch(
  common_name = "yelloweye rockfish", 
  survey = "NWFSC.Combo")

catch_filterFALSE <- pull_catch(
  common_name = "yelloweye rockfish", 
  survey = "NWFSC.Combo",
  standard_filtering = FALSE)

bio = pull_bio(
  common_name = "yelloweye rockfish", 
  survey = "NWFSC.Combo") |>
  dplyr::filter(Latitude_dd >= 42)

bio_nolatcutoff = pull_bio(
  common_name = "yelloweye rockfish", 
  survey = "NWFSC.Combo")

bio_nolatcutoff_filterFALSE = pull_bio(
  common_name = "yelloweye rockfish", 
  survey = "NWFSC.Combo",
  standard_filtering = FALSE)



```


### Length composition data
```{r, warning = FALSE, message = FALSE}

# Use raw length comps, you can tell this is what was used in the previous assessment because whole numbers are used
# The steward-hamel method is used to get the Nsamps, this is better than the simplified stewart-hamel approach used in the 2017 assessment and is fine to use in this years update according to the instructors.
# Nsamp method is the Stewart Hammel method where 
# Nsamp = n_tows + 0.0707 * n_fish when n_fish/n_tows < 55 and
# Nsamp = 4.89 * n_tows when n_fish/n_tows >= 55

raw_length_comps_n_trips <- get_raw_comps(
    data = bio_nolatcutoff,
    comp_bins = seq(10, 74, 2),
    comp_column_name = "length_cm",
    input_n_method = c("tows"),
    two_sex_comps = FALSE,
    fleet = 11,
    month = 7, 
    dir = getwd()#,
   # printfolder = "") 
)

raw_length_comps_n_trips <- raw_length_comps_n_trips$unsexed |>
  dplyr::rename(n_tows = input_n)

raw_length_comps_n_fish <- get_raw_comps(
    data = bio_nolatcutoff,
    comp_bins = seq(10, 74, 2),
    comp_column_name = "length_cm",
    input_n_method = c("total_samples"),
    two_sex_comps = FALSE,
    fleet = 11,
    month = 7#, 
    #dir = getwd()#,
    #printfolder = "")
  )

raw_length_comps_n_fish <- raw_length_comps_n_fish$unsexed |>
  dplyr::rename(n_samps = input_n)|>
  dplyr::select(year, n_samps)

raw_length_comps <- raw_length_comps_n_trips |>
  left_join(raw_length_comps_n_fish, by = "year") |>
  dplyr::mutate(
    n_samps_per_tow = n_samps / n_tows,
    Nsamp = round(ifelse(n_samps_per_tow < 55, (n_tows + 0.0707*n_samps), 4.89*n_tows), digits = 2)) |>
  dplyr::select(year, month, fleet, sex, partition, Nsamp, everything(), -n_samps_per_tow, -n_samps, -n_tows) #|>
  #write.csv(file.path(dirname(dirname(getwd())), "Rcode", "survey_length_age", "length_cm_unsexed_raw_10_74_yelloweye rockfish_groundfish_slope_and_shelf_combination_survey.csv"), row.names = FALSE)


############################################

raw_length_comps_n_trips_filterFALSE <- get_raw_comps(
    data = bio_nolatcutoff_filterFALSE,
    comp_bins = seq(10, 74, 2),
    comp_column_name = "length_cm",
    input_n_method = c("tows"),
    two_sex_comps = FALSE,
    fleet = 11,
    month = 7#, 
    #dir = getwd()#,
   # printfolder = "") 
)

raw_length_comps_n_trips_filterFALSE <- raw_length_comps_n_trips_filterFALSE$unsexed |>
  dplyr::rename(n_tows = input_n)

raw_length_comps_n_fish_filterFALSE <- get_raw_comps(
    data = bio_nolatcutoff_filterFALSE,
    comp_bins = seq(10, 74, 2),
    comp_column_name = "length_cm",
    input_n_method = c("total_samples"),
    two_sex_comps = FALSE,
    fleet = 11,
    month = 7#, 
    #dir = getwd()#,
    #printfolder = "")
  )

raw_length_comps_n_fish_filterFALSE <- raw_length_comps_n_fish_filterFALSE$unsexed |>
  dplyr::rename(n_samps = input_n)|>
  dplyr::select(year, n_samps)

raw_length_comps_filterFALSE <- raw_length_comps_n_trips_filterFALSE |>
  left_join(raw_length_comps_n_fish_filterFALSE, by = "year") |>
  dplyr::mutate(
    n_samps_per_tow = n_samps / n_tows,
    Nsamp = round(ifelse(n_samps_per_tow < 55, (n_tows + 0.0707*n_samps), 4.89*n_tows), digits = 2)) |>
  dplyr::select(year, month, fleet, sex, partition, Nsamp, everything(), -n_samps_per_tow, -n_samps, -n_tows) #|>
  #write.csv(file.path(dirname(dirname(getwd())), "Rcode", "survey_length_age", "length_cm_unsexed_raw_10_74_yelloweye rockfish_groundfish_slope_and_shelf_combination_survey.csv"), row.names = FALSE)

```



### Marginal age composition data
```{r, warning = FALSE, message = FALSE}
# Expanded age comps are not used in the yelloweye assessment
# age_comps <- get_expanded_comps(
#     bio_data = bio,
#     catch_data = catch,
#     comp_bins = 0:65,
#     strata = WCGBTS_strata,
#     comp_column_name = "age",
#     output = "full_expansion_ss3_format",
#     two_sex_comps = FALSE, #single sex model
#     input_n_method = "stewart_hamel")
# 
# plot_comps(
#   data = age_comps)

raw_age_comps_n_trips <- get_raw_comps(
    data = bio_nolatcutoff,
    comp_bins = 0:65,
    input_n_method = "tows",
    comp_column_name = "age",
    fleet = 11,
    ageerr = 2,
    month = 7,
    #dir = getwd(),
    #printfolder = "", 
    two_sex_comps = FALSE) 

raw_age_comps_n_trips <- raw_age_comps_n_trips$unsexed |>
  dplyr::rename(n_tows = input_n)

raw_age_comps_n_fish <- get_raw_comps(
    data = bio_nolatcutoff,
    comp_bins = 0:65,
    input_n_method = "total_samples",
    comp_column_name = "age",
    fleet = 11,
    ageerr = 2,
    month = 7,
    #dir = getwd(),
    #printfolder = "", 
    two_sex_comps = FALSE)

raw_age_comps_n_fish <- raw_age_comps_n_fish$unsexed |>
  dplyr::rename(n_samps = input_n)|>
  dplyr::select(year, n_samps)

raw_age_comps <- raw_age_comps_n_trips |>
  left_join(raw_age_comps_n_fish, by = "year") |>
  mutate(
    n_samps_per_tow = n_samps / n_tows,
    Nsamp = round(ifelse(n_samps_per_tow < 55, (n_tows + 0.0707 * n_samps), 4.89*n_tows), digits = 2)) |>
  dplyr::select(year, month, fleet, sex, partition, Nsamp, everything(), -n_samps_per_tow, -n_samps, -n_tows) #|>
  #write.csv(file.path(dirname(dirname(getwd())), "Rcode", "survey_length_age", "age_unsexed_raw_0_65_yelloweye rockfish_groundfish_slope_and_shelf_combination_survey.csv"), row.names = FALSE)

# plot_comps(
#   data = raw_age_comps_n_trips)


######################################################

raw_age_comps_n_trips_filterFALSE <- get_raw_comps(
    data = bio_nolatcutoff_filterFALSE,
    comp_bins = 0:65,
    input_n_method = "tows",
    comp_column_name = "age",
    fleet = 11,
    ageerr = 2,
    month = 7,
    #dir = getwd(),
    #printfolder = "", 
    two_sex_comps = FALSE) 

raw_age_comps_n_trips_filterFALSE <- raw_age_comps_n_trips_filterFALSE$unsexed |>
  dplyr::rename(n_tows = input_n)

raw_age_comps_n_fish_filterFALSE <- get_raw_comps(
    data = bio_nolatcutoff_filterFALSE,
    comp_bins = 0:65,
    input_n_method = "total_samples",
    comp_column_name = "age",
    fleet = 11,
    ageerr = 2,
    month = 7,
    #dir = getwd(),
    #printfolder = "", 
    two_sex_comps = FALSE)

raw_age_comps_n_fish_filterFALSE <- raw_age_comps_n_fish_filterFALSE$unsexed |>
  dplyr::rename(n_samps = input_n)|>
  dplyr::select(year, n_samps)

raw_age_comps_filterFALSE <- raw_age_comps_n_trips_filterFALSE |>
  left_join(raw_age_comps_n_fish_filterFALSE, by = "year") |>
  mutate(
    n_samps_per_tow = n_samps / n_tows,
    Nsamp = round(ifelse(n_samps_per_tow < 55, (n_tows + 0.0707 * n_samps), 4.89*n_tows), digits = 2)) |>
  dplyr::select(year, month, fleet, sex, partition, Nsamp, everything(), -n_samps_per_tow, -n_samps, -n_tows) #|>
  #write.csv(file.path(dirname(dirname(getwd())), "Rcode", "survey_length_age", "age_unsexed_raw_0_65_yelloweye rockfish_groundfish_slope_and_shelf_combination_survey.csv"), row.names = FALSE)


```


### Conditional age-at-length data
```{r, warning = FALSE, message = FALSE}
# The previous function used was actually deprecated and this function should have been used.
caal <- get_raw_caal(
  dat = bio_nolatcutoff, 
  len_bins = seq(10, 74, 2), 
  age_bins = 0:65#,
  #dir = file.path(dirname(dirname(getwd())), "Data", "processed", "NWFSC.Combo_CAAL"),
  #printfolder = "")
)

# Reformat data to only have one sex
one_sex_caal <- caal |>
  dplyr::arrange(year, Lbin_lo) |>
  tidyr::pivot_longer(cols = f0:m65, names_to = "age", values_to = "count") |>
  dplyr::mutate(input_n = case_when(
    grepl("m", age) ~ 0,
    grepl("f", age) ~ input_n
  )) |>
  mutate(age = as.numeric(stringr::str_remove_all(age, "f|m"))) |>
  dplyr::group_by(year, Lbin_lo, age) |>
  dplyr::summarize(month = 7,
                   fleet = 11,
                   sex = 0,
                   partition = unique(partition),
                   ageerr = 2,
                   Lbin_hi = unique(Lbin_hi),
                   Nsamp = sum(input_n),
                   count = sum(count)) |>
  dplyr::select(year, month, fleet, sex, partition, ageerr, Lbin_lo, Lbin_hi, Nsamp, age, count) |>
  tidyr::pivot_wider(names_from = "age", values_from = "count")

#write.csv(one_sex_caal, file = file.path(dirname(dirname(getwd())), "Data", "processed", "NWFSC.Combo_CAAL", "processed_one_sex_caal.csv"), row.names = FALSE)

# PlotMap.fn(
#   dat = catch)


#########################################

caal_filterFALSE <- get_raw_caal(
  dat = bio_nolatcutoff_filterFALSE, 
  len_bins = seq(10, 74, 2), 
  age_bins = 0:65#,
  #dir = file.path(dirname(dirname(getwd())), "Data", "processed", "NWFSC.Combo_CAAL"),
 # printfolder = "")
)

# Reformat data to only have one sex
one_sex_caal_filterFALSE <- caal_filterFALSE |>
  dplyr::arrange(year, Lbin_lo) |>
  tidyr::pivot_longer(cols = f0:m65, names_to = "age", values_to = "count") |>
  dplyr::mutate(input_n = case_when(
    grepl("m", age) ~ 0,
    grepl("f", age) ~ input_n
  )) |>
  mutate(age = as.numeric(stringr::str_remove_all(age, "f|m"))) |>
  dplyr::group_by(year, Lbin_lo, age) |>
  dplyr::summarize(month = 7,
                   fleet = 11,
                   sex = 0,
                   partition = unique(partition),
                   ageerr = 2,
                   Lbin_hi = unique(Lbin_hi),
                   Nsamp = sum(input_n),
                   count = sum(count)) |>
  dplyr::select(year, month, fleet, sex, partition, ageerr, Lbin_lo, Lbin_hi, Nsamp, age, count) |>
  tidyr::pivot_wider(names_from = "age", values_from = "count")





```

### Weight-Length Relationship
```{r, warning = FALSE, message = FALSE}
# Ian said we can just use the nwfsc.combo data and don't need to filter by the same latitudes as the index was, just use all data
wt_len_est <- estimate_weight_length(
  bio_nolatcutoff,
  col_length = "Length_cm",
  col_weight = "Weight_kg",
  verbose = FALSE
)

wt_len_est

plot_weight_length(
  bio_nolatcutoff,
  dir = file.path(getwd(), "filter_TF_test"),
  estimates = wt_len_est,
  col_length = "length_cm",
  col_weight = "weight_kg",
  two_sex = FALSE,
  add_save_name = "filterTRUE",
  height = 7,
  width = 7,
  dpi = 300
)



######################################
wt_len_est_filterFALSE <- estimate_weight_length(
  bio_nolatcutoff_filterFALSE,
  col_length = "Length_cm",
  col_weight = "Weight_kg",
  verbose = FALSE
)

wt_len_est_filterFALSE

plot_weight_length(
  bio_nolatcutoff_filterFALSE,
  dir = file.path(getwd(), "filter_TF_test"),
  estimates = wt_len_est_filterFALSE,
  col_length = "length_cm",
  col_weight = "weight_kg",
  two_sex = FALSE,
  add_save_name = "filterFALSE",
  height = 7,
  width = 7,
  dpi = 300
)




```


## Composition comparison plots
### NWFSC.Combo Comparisons
```{r}
# Lengths
nwfsc_lengths_filterTRUE <- raw_length_comps|>
  mutate(filter = "TRUE") |>
  dplyr::select(year, filter, `u10`:`u74`) |>
  tidyr::pivot_longer(cols = c(-year, -filter), names_to = "length", values_to = "freq") |>
  dplyr::mutate(length = gsub("u", "", length)) |>
  dplyr::group_by(year) |>
  dplyr::mutate(
    freq = freq / sum(freq),
    length = as.numeric(length)
  )

nwfsc_lengths_filterFALSE <- raw_length_comps_filterFALSE|>
  mutate(filter = "FALSE") |>
  dplyr::select(year, filter, `u10`:`u74`) |>
  tidyr::pivot_longer(cols = c(-year, -filter), names_to = "length", values_to = "freq") |>
  dplyr::mutate(length = gsub("u", "", length)) |>
  dplyr::group_by(year) |>
  dplyr::mutate(
    freq = freq / sum(freq),
    length = as.numeric(length)
  )

length_together <- rbind(nwfsc_lengths_filterTRUE, nwfsc_lengths_filterFALSE)
comparison_plot_length <- length_together |>
  dplyr::filter(freq > 0) |>
  ggplot2::ggplot(aes(x = year, y = length, col = filter, size = freq)) +
  ggplot2::geom_point(position = position_dodge(0.5))
ggsave(plot = comparison_plot_length, "nwfsc_length_comp_TF_comparisons.png", path = file.path(getwd(), "filter_TF_test", "plots"))

# Ages
nwfsc_ages_filterTRUE <- raw_age_comps |>
  mutate(filter = "TRUE") |>
  dplyr::select(year, filter, `u0`:`u65`) |>
  tidyr::pivot_longer(cols = c(-year, -filter), names_to = "age", values_to = "freq") |>
  dplyr::mutate(age = gsub("u", "", age)) |>
  dplyr::group_by(year) |>
  dplyr::mutate(
    freq = freq / sum(freq),
    # NsampsFreq = freq/unique(Nsamps), still waiting on these
    age = as.numeric(age)
  )

nwfsc_ages_filterFALSE <- raw_age_comps_filterFALSE |>
  mutate(filter = "FALSE") |>
  dplyr::select(year, filter, `u0`:`u65`) |>
  tidyr::pivot_longer(cols = c(-year, -filter), names_to = "age", values_to = "freq") |>
  dplyr::mutate(age = gsub("u", "", age)) |>
  dplyr::group_by(year) |>
  dplyr::mutate(
    freq = freq / sum(freq),
    # NsampsFreq = freq/unique(Nsamps), still waiting on these
    age = as.numeric(age)
  )

age_together <- rbind(nwfsc_ages_filterTRUE, nwfsc_ages_filterFALSE)
comparison_plot_age <- age_together |>
  dplyr::filter(freq > 0) |>
  ggplot2::ggplot(aes(x = year, y = age, col = filter, size = freq)) +
  ggplot2::geom_point(position = position_dodge(0.5))
ggsave(plot = comparison_plot_age, "nwfsc_age_comp_TF_comparisons.png", path = file.path(getwd(), "filter_TF_test", "plots"))
#ggsave(plot = comparison_plot, "nwfsc_age_comp_comparisons.png", path = file.path(getwd(), "Rcode", "survey_length_age", "NWFSC.Combo_plots"))
```
